Holomorphic field theories and higher algebra

### 2024-08-19 draft.tex
merge

A theory is usually described, in part, in terms of a space of fields which comprise some possibly nonlinear space like a mapping space or the space of connections. 
In perturbative field theory, which describes the formal neighborhood of the theory at points in such spaces, the fields $\Phi$ are typically described by the smooth sections of a bundle on spacetime.
The equations of motion which dictate the dynamics of the particular model are encoded by a partial differential equation heuristically of the form 
\[
\sum_{k \geq 1} D_k(\Phi,\ldots,\Phi) = 0 ,
\]
where $D_k$ is a $k$-ary polydifferential operator.
As an example of a Euclidean theory, the $\phi^4$-theory on a Riemannian four-manifold $M$ is described by the equation
\[
\Delta \phi + \frac{1}{3!} \phi^3 = 0 
\]
where $\Phi = \phi$ is simply a smooth function on $M$.

In a {\em holomorphic theory} one starts with a complex manifold $X$ as the spacetime.
This means, in part, that the complexified tangent bundle of $X$ decomposes as a sum of vector bundles $\T_X \oplus \Bar{\T}_X$ where $\T_X$ (respectively $\Bar{T}_X$) is the holomorphic (respectively anti-holomorphic) tangent bundle.
In local coordinates sections of $\T_X$ have the form $f^i \del / \del z_i$ whereas sections of $\Bar{\T}_X$ have the form $f^i \del / \del \zbar_i$.

Globally, holomorphic theories are described like in the usual case by mapping spaces or spaces of connections.
The key feature in a holomorphic theory is that the equations of motion dictate that fields be holomorphic.
Perturbatively, where fields are sections of some holomorphic vector bundle $V$, we can make this precise. 
Prescribing a holomorphic structure on a vector bundle $V$ is equivalent to the data of a $\dbar$-operator
\beqn
\dbar \colon \Omega^{0,q}(X, V) \to \Omega^{0,q+1}(X,V) 
\eeqn
where $\Omega^{0,q}(X,V)$ is the space of smooth sections of the vector bundle $\wedge^q\Bar{T}_X^* \otimes V$. 
Holomorphicity of $V$ boils down to the requirement that $\dbar \circ \dbar = 0$.
The Dolbeault complex of $V$ is the resulting complex $(\Omega^{0,\bu}(X, V), \dbar)$ where $\Omega^{0,q}(X,V)$ sits in degree $+q$.
A standard result in complex geometry is that the Dolbeault complex is a resolution for the sheaf of holomorphic sections of the holomorphic vector bundle $V$.
In the case that $V$ is the trivial bundle $\ul{\CC}$ the Dolbeault complex $\Omega^{0,\bu}(X)$ is a resolution for the sheaf of holomorphic functions on $X$.

The fields of a holomorphic theory are given by elements of the Dolbeault complex $\Omega^{0,\bu}(X,V)$.
The equations of motion in a holomorphic theory take the form
\beqn
\dbar \Phi + \sum_{k \geq 1} D_k(\Phi,\ldots,\Phi) = 0 ,
\eeqn
where $D_k$ are $k$-ary {\em holomorphic} polydifferential operators.
This means that in local coordinates $D_k$ only depend on the holomorphic coordinates $\{z_i\}$ and involve only holomorphic derivatives $\del / \del z_i$.

@@

Let us begin with the example of holomorphic Chern--Simons theory, 
which we have already said describes the moduli space of Maurer--Cartan elements describe deformations of the trivial holomorphic $G$-bundle. 
When $X$ is a Calabi--Yau three-fold and $\fg$ is equipped with an invariant pairing $\<\cdot , \cdot \>$, 
this formal moduli space has a very special property: 
it arises from the variational problem of an action functional, namely
\beqn
S = \int_X \Omega \wedge {\rm CS}(A) 
\eeqn
where $\Omega$ is the holomorphic volume form and 
\beqn
{\rm CS}(A) = \frac12 \< A ,\d A \> + \frac16 \<A , [A,A]\>
\eeqn
is the Chern--Simons Lagrangian for the connection $(0,1)$-form $A$.
Notice that due to the presence of the holomorphic volume form $\Omega$ on the $\dbar$ operator contributes to the kinetic term and so the action can be written as 
\beqn
S = \int_X \Omega \wedge \left(\frac12 \< A ,\dbar A \> + \frac16 \<A , [A,A]\>\right) .
\eeqn

@@

\subsubsection{Holomorphic scalar field theories}

A typical scalar field theory is the free boson which exists on any Riemannian manifold $\Sigma$.
Its action functional is simply $\int_\Sigma \phi \triangle \phi$ where $\phi$ is a smooth function on $\Sigma$.
When $\Sigma$ is a surface this theory can be written in such a way that makes its complex structure apparent.
First, we complexify the field $\phi$ which we can view as a pair of fields $\gamma, \Bar{\gamma}$.
When we do this, we can split it up into its chiral and anti-chiral `sectors'.
By chiral (anti-chiral) we mean part involving $\gamma$ ($\Bar{\gamma}$).
The holomorphic part is described by the following action
\beqn
\int_\Sigma \beta \wedge \dbar \gamma
\eeqn
where $\beta$ is a $(1,0)$ form on $\Sigma$ and should be thought of as a Lagrange multiplier for the field $\del \Bar{\gamma}$ which we now think about as having its own degrees of freedom (this is called the first-order formalism).
There is also a coupling between the chiral and anti-chiral sectors, 
for more details we refer to \cite{Kapustin,GGW}. 
This piece of the free boson is the simplest holomorphic field theory; aptly called the $\beta-\gamma$-\textit{system}.
It plays a role in superstring theory as the free field theory underlying bosonic ghosts corresponding to worldsheet supersymmetry~\cite{Polchinski}.

There is a holomorphic field theory defined on an arbitrary complex manifold as a simple generalization of this `$\beta-\gamma$ system'.
The input data is simply a complex manifold $X$ of arbitrary complex dimension $n$ together with a (super) holomorphic vector bundle $V$ on~$X$.
Above we took this bundle to be trivial.
The fields of the theory consist of pairs 
\begin{align*}
\gamma & \in \Gamma(X,V)\\
\beta & \in \Omega^{n,n-1}(X,V^*) .
\end{align*}
The action functional is identical $\int_X \beta \wedge \dbar \gamma$.

Notice that when $n = \dim_\C X > 1$ that there is a linear symmetry present in the above action given by shifting $\beta$ by a $\dbar$-exact form
\beqn
\beta \mapsto \beta + \dbar \eta , \quad \eta \in \Omega^{n,n-2}(X,V^*)
\eeqn
For this model to be well-posed at the quantum level we must include the ghosts $\eta$.
Further, $\eta$ can be shifted by an exact form in a way that does not effect the above gauge transformation, so we must also include ghosts for ghosts, etc..
This is one of the key differences between the ordinary $\beta-\gamma$ system and its higher dimensional version.
Nevertheless, many techniques we have for understanding free field theories can still by applied to this model as we will explain later in this note.

@@

@@

If $X,Y$ are any two complex manifolds there is a theory whose fields consist, in part, of the derived moduli space of smooth maps $\RR {\rm Map}(X,Y)$ and whose equations of motion dictate that such maps be holomorphic.
This is a nonlinear generalization of the $\beta\gamma$ system we just introduced.

The precise space of fields, in cohomological degree zero, consist of pairs $(\gamma, \beta)$ where $\gamma \colon X \to Y$ is a smooth map and 
\[
\beta \in \Omega^{n,n-1}(X, f^* \T^*_Y) 
\]
where $\T^*_Y$ is the holomorphic cotangent bundle of $Y$. 
The action functional is the familiar one
\beqn
\int_X \beta \dbar \gamma .
\eeqn 
Solutions to the equations of motion describe, in part, holomorphic maps $X \to Y$, but there is also derived information contained in this model.

@@

%In complex dimension $3$ there is a related theory, called Kodaira--Spencer theory, which only consists of the field $\mu$ describing the complex structure in the above example.
Let $X$ be a complex threefold equipped with a Calabi--Yau structure; we denote by $\Omega \in \Omega^{3,hol}(X)$ the holomorphic volume form.
The primary field of Kodaira spencer theory is a $(0,1)$ form on $X$ with values in the holomorphic tangent bundle $\T_X$
\beqn
\mu \in \Omega^{0,1}(X,\T_X) .
\eeqn 
A crucial assumption of this model is that we must assume that $\mu$ preserve the holomorphic volume form $\Omega$, meaning its divergence vanishes ${\rm div}_\Omega \mu = 0$.
The action functional is very similar to the Chern--Simons action
\beqn
S(\mu) = \int_X \left(\frac12 \mu \dbar \del^{-1} \mu + \frac13 \mu^3\right) .
\eeqn
In the integrand we use the Calabi--Yau form to obtain a $(3,3)$-form.
The integrand is not defined globally as it involves the distributional operator $\del^{-1}$.
Locally it is well-defined as any divergence-free vector field on affine space can be written as the divergence of a bivector field.
Nevertheless, this model can be carefully put into the rigorous framework of the Batalin--Vilkovisky formalism \cite{CLbcov1}.
Forgoing the apparent global issues, we can still apply formal variational calculus to the action to obtain the equation of motion
\beqn
\dbar \mu + \frac12 [\mu,\mu] = 0 ,
\eeqn
where $[-,-]$ is the Schouten--Nijenhuis bracket (which reduces to just the natural bracket when restricted to vector fields).
The reader can recognize this equation as the Maurer--Cartan equation controlling deformations of complex structure $\dbar\rightsquigarrow \dbar + \mu$.
\subsection{Local version and examples from supersymmetric field theories}

>>>>

A theory is usually described, in part, in terms of a space of fields which comprise some possibly nonlinear space like a mapping space or the space of connections. 
In perturbative field theory, which describes the formal neighborhood of the theory at points in such spaces, the fields $\Phi$ are typically described by the smooth sections of a vector bundle on spacetime.
The equations of motion which dictate the dynamics of the particular model are encoded by a series of partial differential equations of the form 
\[
\sum_{k \geq 1} D_k(\Phi,\ldots,\Phi) = 0 ,
\]
where $D_k$ is a $k$-ary polydifferential operator.
As an example, the scalar $\phi^4$-theory on a Riemannian $M$ is described by the equation
\[
\Delta \phi + \frac{1}{3!} \phi^3 = 0 
\]
where $\Phi = \phi$ is simply a smooth function on $M$.
These equations of motion are natural to the Riemannian geometry of the manifold $M$, and can be derived from the Euler--Lagrange equations of the simple action functional $\int_M \phi \Delta \phi \, \op{dvol}$.

As another example, consider perturbative Chern--Simons theory whose fields are a principal $G$-bundle $P$ on a smooth three-manifold $M$ equipped with a connection.
The equations of motion is the condition that the connection is flat 
\beqn
F_\nabla = 0 .
\eeqn
Perturbatively, near the trivial flat connection, the fields are simply $\lie{g} = \op{Lie}(G)$-valued one-forms $A \in \Omega^1(M, \lie{g})$.
The condition that perturbed connection $\d + A$ is flat is the Maurer--Cartan equation
\beqn
\d A + \frac12 [A,A] = 0.
\eeqn
These equations of motion are topological in the sense that they only depend on the smooth manifold~$M$.
When $\lie{g}$ is equipped with a non-degenerate, invariant, bilinear form $\kappa$ (for example, if $G$ is semi-simple there is a standard choice) then these equations of motion can be derived from the famous Chern--Simons action functional $\int_M CS(A)$ where
\beqn\label{eqn:CS}
CS(A) = \frac12 \kappa(A, \d A) + \frac16 \kappa(A, [A,A]) .
\eeqn

In a {\em holomorphic theory} one starts with a complex manifold $X$ as the spacetime.
%This means, in part, that the complexified tangent bundle of $X$ decomposes as a sum of vector bundles $\T_X \oplus \Bar{\T}_X$ where $\T_X$ (respectively $\Bar{T}_X$) is the holomorphic (respectively anti-holomorphic) tangent bundle.
%In local coordinates sections of $\T_X$ have the form $f^i \del / \del z_i$ whereas sections of $\Bar{\T}_X$ have the form $f^i \del / \del \zbar_i$. 
Globally, holomorphic theories are described like in the usual case by mapping spaces or spaces of connections.
The key feature in a holomorphic theory is that the equations of motion dictate that fields be holomorphic in an appropriate sense.
Perturbatively, where fields are sections of some (possibly graded) holomorphic vector bundle $V$, we can make this precise. 
Prescribing a holomorphic structure on a vector bundle $V$ is equivalent to the data of a $\dbar$-operator
\beqn
\dbar \colon \Omega^{0,q}(X, V) \to \Omega^{0,q+1}(X,V) 
\eeqn
where $\Omega^{0,q}(X,V)$ is the space of smooth sections of the vector bundle $\wedge^q\Bar{T}_X^* \otimes V$. 
Holomorphicity of $V$ boils down to the requirement that $\dbar \circ \dbar = 0$.
The Dolbeault complex of $V$ is
\beqn
\left(\Omega^{0,\bu}(X, V), \dbar \right) = \left(\oplus_{q \geq 0} \Omega^{0,q}(X, V)[-q], \dbar \right)
\eeqn
where (in the case that $V$ sits in degree zero) the space $\Omega^{0,q}(X,V)$ sits in degree $+q$.
A standard result in complex geometry is that the Dolbeault complex is a resolution for the sheaf of holomorphic sections of the holomorphic vector bundle $V$.
%In the case that $V$ is the trivial bundle $\ul{\CC}$ the Dolbeault complex $\Omega^{0,\bu}(X)$ is a resolution for the sheaf of holomorphic functions on $X$.

The fields of a holomorphic theory are given by elements of the Dolbeault complex $\Omega^{0,q}(X,V)$ for some $q$.
The equations of motion in a holomorphic theory take the form
\beqn
\dbar \Phi + \sum_{k \geq 1} D_k(\Phi,\ldots,\Phi) = 0 ,
\eeqn
where $D_k$ are $k$-ary {\em holomorphic} polydifferential operators.
This means that in local coordinates $D_k$ only depend on the holomorphic coordinates $\{z_i\}$ and involve only holomorphic derivatives $\del / \del z_i$.

For a simple example, consider the following holomorphic analog of Chern--Simons theory.
The fields are principal $G$-bundles on a three-dimensional \textit{complex} manifold $X$, and the equations of motion give the condition that the principal $G$-bundle is holomorphic.
Perturbatively, near the trivial $G$-bundle, the fields of this holomorphic theory are $\lie{g}$-valued $(0,1)$-forms $A \in \Omega^{0,1}(X, \lie{g})$.
The condition that $\d + A$ prescribes a holomorphic structure is again a Maurer--Cartan equation
\beqn
\dbar A + \frac12 [A,A] = 0 .
\eeqn
Like ordinary Chern--Simons theory, in order for this equation to arise from a principal of least action one needs to fix a non-degenerate, invariant, bilinear form $\kappa$ on $\lie{g}$.
Additionally, one needs to fix a holomorphic volume form $\Omega$ on $X$; in other words, $X$ is a Calabi--Yau manifold.
Then, the equations of motion is precisely the Euler--Lagrange equation for the action
\beqn
S_{hCS} (A) = \int_X \Omega \wedge CS(A) 
\eeqn
where $CS(A)$ is exactly as in \eqref{eqn:CS}.\footnote{Notice that due to the presence of the holomorphic volume form $\Omega$ only the $\dbar$ operator contributes to the kinetic term and so the action can be written as $\int_X \Omega \wedge \left(\frac12 \kappa( A ,\dbar A ) + \frac16 \kappa(A , [A,A]) \right)$.}

@@

@@

@@

Holomorphic BF theory is easy to formulate classically, but it can admit peculiar properties at the quantum level.
For example, in \cite{semichiral} it is shown that on $X = \C^2$ the quantization of holomorphic BF theory is almost topological.\footnote{More precisely, the divergence-free part of the stress tensor is rendered cohomologically trivial to first-order in perturbation theory.}

@@

If $X,Y$ are any two complex manifolds there is a theory whose fields consist, in part, of the derived moduli space of smooth maps $\RR {\rm Map}(X,Y)$ and whose equations of motion dictate that such maps be holomorphic.

This class of holomorphic field theories are simple higher-dimensional generalizations of the well-studied `$\beta-\gamma$ system' which appears in: superstring theory~\cite{Polchinski}, as the twist (see below)of the $\cN=(2,0)$ supersymmetric $\sigma$-model, and as the chiral part of the scalar field theory on a Riemann surface.

Consider the case that $Y$ is simply $\C$ and suppose that $X$ is $n$-dimensional.
The fields of the model are
\begin{align*}
\gamma & \in \Gamma(X)\\
\beta & \in \Omega^{n,n-1}(X) .
\end{align*}
and the action is
\beqn
\int_\Sigma \beta \wedge \dbar \gamma .
\eeqn
There is a peculiarity of this theory when $n = \dim_\C X > 1$: in the case, there is a linear symmetry by shifting $\beta$ by a $\dbar$-exact form
\beqn
\beta \mapsto \beta + \dbar \eta , \quad \eta \in \Omega^{n,n-2}(X,V^*)
\eeqn
For this model to be well-posed at the quantum level we must include the ghosts $\eta$.
Further, $\eta$ can be shifted by an exact form in a way that does not effect the above gauge transformation, so we must also include ghosts for ghosts, etc..
This is one of the key differences between the ordinary $\beta-\gamma$ system and its higher dimensional version.
Nevertheless, many techniques we have for understanding free field theories can still be applied to this model.

There is the following nonlinear generalization of the $\beta\gamma$ system when $Y$ is an arbitrary complex manifold.
The space of fields consist of pairs $(\gamma, \beta)$ where $\gamma \colon X \to Y$ is a smooth map and 
\[
\beta \in \Omega^{n,n-1}(X, f^* \T^*_Y) 
\]
where $\T^*_Y$ is the holomorphic cotangent bundle of $Y$. 
The action functional is the familiar one
\beqn
\int_X \beta \dbar \gamma .
\eeqn 
Solutions to the equations of motion describe, in part, holomorphic maps $X \to Y$, but there is also derived information contained in this model.
A detailed formulation of this theory has been developed in \cite{BWthesis}.

@@

%In complex dimension $3$ there is a related theory, called Kodaira--Spencer theory, which only consists of the field $\mu$ describing the complex structure in the above example.
Let $X$ be a complex threefold equipped with a Calabi--Yau structure; we denote by $\Omega \in \Omega^{3,hol}(X)$ the holomorphic volume form.
The primary field of Kodaira spencer theory is a $(0,1)$ form on $X$ with values in the holomorphic tangent bundle $\T_X$
\beqn
\mu \in \Omega^{0,1}(X,\T_X) .
\eeqn 
A crucial assumption of this model is that we must assume that $\mu$ preserve the holomorphic volume form $\Omega$, meaning its divergence vanishes ${\rm div}_\Omega \mu = 0$.
The action functional is very similar to the Chern--Simons action
\beqn\label{eqn:KS1}
S(\mu) = \int_X \left(\frac12 \mu \dbar \del^{-1} \mu + \frac13 \mu^3\right) .
\eeqn
In the integrand we use the Calabi--Yau form to obtain a $(3,3)$-form.
The integrand is not defined globally as it involves the distributional operator $\del^{-1}$.
Locally it is well-defined as any divergence-free vector field on affine space can be written as the divergence of a bivector field.
Nevertheless, this model can be carefully put into the rigorous framework of the Batalin--Vilkovisky formalism \cite{CLbcov1}.
Forgoing the apparent global issues, we can still apply formal variational calculus to the action to obtain the equation of motion
\beqn
\dbar \mu + \frac12 [\mu,\mu] = 0 ,
\eeqn
where $[-,-]$ is the Schouten--Nijenhuis bracket (which reduces to just the natural bracket when restricted to vector fields).
The reader can recognize this equation as the Maurer--Cartan equation controlling deformations of complex structure $\dbar\rightsquigarrow \dbar + \mu$.

\subsubsection{Holomorphic scalar field theories / $\sigma$-models}

A typical scalar field theory is the free boson which exists on any Riemannian manifold $\Sigma$.
Its action functional is simply $\int_\Sigma \phi \triangle \phi$ where $\phi$ is a smooth function on $\Sigma$.
We saved this example for last since its holomorphic analog is a bit more elusive than the previous examples.

We start with complex dimension $1$.
The usual scalar field theory can be written as $\int_{\Sigma} \phi \del \dbar \phi$, where $\Sigma$ is a Riemann surface.
This theory is not holomorphic, but we can force it to be by demanding the constraint $\dbar \phi = 0$.
Importantly, this constraint is \textit{not} imposed by an equation of motion.
When we do this, we can express the action functional in terms of the $(1,0)$-form $\alpha \in \Omega^{1,0}(\Sigma)$ as
\beqn\label{eqn:scalar1}
\int \alpha \dbar \del^{-1} \alpha .
\eeqn
This is formally similar to the kinetic term in the action functional of Kodaira--Spencer theory \eqref{eqn:KS1}.

There is a higher-dimensional analog of the chiral sector of the scalar field.
Suppose that $X$ is a complex manifold of \textit{odd} dimension $2n+1$.
The fields are certain $(n+1,n)$-forms $\alpha \in \Omega^{n+1,n}(X)$ which satisfy the constraint
\beqn
\del \alpha = 0.
\eeqn
Note that when $n=0$, so $X = \Sigma$ is simply a Riemann surface, this constraint is automatic.
The action functional is the same \eqref{eqn:scalar1}.
When $n=1$ this is identical to the kinetic part of the Kodaira--Spencer action.\footnote{Indeed, when $X$ is a Calabi--Yau three-fold there is an isomorphism $\Omega^{2,1}(X) \simeq \Omega^{0,1}(X, \T_X)$.}
Generally, this theory is the holomorphic part of a higher-form scalar theory, with action
\beqn
\int_X \|C\|^2 \op{dvol} = \int_X C \d \star \d C , \quad C \in \Omega^{2n+1}(X) .
\eeqn
For more details we refer to \cite{GRWwzw}.

\subsection{Local version and examples from supersymmetric field theories}
